---
title: "exercise-11"
author: "Yen Do"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(skimr)
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f)
skim(d)
```

Categorical variables: order, family, genus, species. Numeric variables: mass(g), gestation(mo), newborn(g), weaning(mo), wean mass(g), AFR(mo), max. life(mo), litter size, litters/year, refs.

### Step 1: Replace all values of -999 (the authors’ code for missing data) with NA.

```{r}
library(naniar)
replace_with_na_all(data = d, condition = ~.x == -999)
```

### Step 2: Drop the variables litter size and refs.

```{r}
library(dplyr)
d <- d |> dplyr::select(-"litter size", - "refs")
```

### Step 3: Log transform all of the other numeric variables.

```{r}
d <- d |> mutate(across(where(is.numeric), log))
head(d)
```

### Step 4: Regression

```{r}
relGest <- lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action=na.exclude)
relWean <- lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action=na.exclude)
relAFR <- lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action=na.exclude)
relLife <- lm(`max. life(mo)` ~ `mass(g)`, data = d, na.action=na.exclude)
relNewbornMass <- lm(`newborn(g)` ~ `mass(g)`, data = d, na.action=na.exclude)
relWeaningMass <- lm(`wean mass(g)` ~ `mass(g)`, data = d, na.action=na.exclude)


d <- d |> mutate(relGest = residuals(relGest),
                 relWean = residuals(relWean),
                 relAFR = residuals(relAFR),
                 relLife = residuals(relLife),
                 relNewbornMass = residuals(relNewbornMass),
                 relWeaningMass = residuals(relWeaningMass))

```

### Step 5: Plot residuals

```{r}
# Plot helper function
plot_residuals <- function(resid_col, title) {
  ggplot(d, aes(x = order, y = .data[[resid_col]])) +
    geom_boxplot() +
    coord_flip() +
    theme_minimal() +
    labs(title = title, y = "Residual", x = "Order")
}

plot_residuals("relLife", "Residual Max Lifespan by Order")
plot_residuals("relNewbornMass", "Residual Newborn Mass by Order")
plot_residuals("relWeaningMass", "Residual Weaning Mass by Order")

```

Which mammalian orders have the highest residual lifespan?

Primates have the highest residual lifespan.

Which mammalian orders have the have highest residual newborn mass?

Macroscelidea have highest residual newborn mass.

Which mammalian orders have the have highest residual weaning mass?

Perissodactya have the have highest residual weaning mass.

### Step 6: Model

```{r}
library(MASS)
library(AICcmodavg)
library(MuMIn)
# Select variables for model
d1 <- d %>%
  drop_na(`max. life(mo)`, `gestation(mo)`, `newborn(g)`,
          `weaning(mo)`, `wean mass(g)`, `litters/year`, `mass(g)`, `AFR(mo)`)

# Full model
m_max_life <- lm(`max. life(mo)` ~ `gestation(mo)` + `newborn(g)` +
                   `weaning(mo)` + `wean mass(g)` +
                   `litters/year` + `mass(g)`, data = d1, na.action = "na.fail")


m_AFR <- lm(`AFR(mo)` ~ `gestation(mo)` + `newborn(g)` +
                   `weaning(mo)` + `wean mass(g)` +
                   `litters/year` + `mass(g)`, data = d1, na.action = "na.fail")

```

#### Model for max. life (mo)

```{r}
# Model for max. life(mo)
mods_life <- dredge(m_max_life, beta = "none", m.lim = c(0, 4))
# best model
mods_life[1, ]

# Number of models have delta AIC < 4
sum(mods_life$delta < 4)

# Calculate and plot the model-averaged coefficients and their CIs for max_life(mo)
(mods.list_life <- get.models(mods_life, subset = TRUE))
coef(mods_life)
(mods.avg_life <- summary(model.avg(mods_life, subset = delta < 4, fit = TRUE)))
confint(mods.avg_life)
plot(mods.avg_life, full = TRUE, intercept = FALSE)

```

Best model (model 40): max. life(mo) \~ gestation(mo) + weaning(mo) + litters/year + mass(g)

2 models have delta AICc \< 4

Model-averaged coefficients: Mammals with longer gestation, weaning period, or higher mass tend to live longer.

wean mass(g) is also positively associated, though its effect is smaller.

litters/year is negatively associated — species with more litters per year tend to have shorter maximum lifespans (consistent with a "fast" life history strategy).

#### Model for AFR(mo)

```{r}
mods_AFR <- dredge(m_AFR, beta = "none", m.lim = c(0, 4))
# best model
mods_AFR[1, ]

# Number of models have delta AIC < 4
sum(mods_AFR$delta < 4)


# Calculate and plot the model-averaged coefficients and their CIs for AFR(mo)
(mods.list_AFR <- get.models(mods_AFR, subset = TRUE))
coef(mods_AFR)
(mods.avg_AFR <- summary(model.avg(mods_AFR, subset = delta < 4, fit = TRUE)))
confint(mods.avg_AFR)
plot(mods.avg_AFR, full = TRUE, intercept = FALSE)


```

Best model (AICc = 321.23): AFR(mo) \~ gestation(mo) + weaning(mo) + litters/year + mass(g)

7 models have delta AICc \< 4

Model-averaged coefficients:

Longer gestation → later reproduction

Fewer litters per year → later reproduction

Longer weaning duration also has a smaller but significant effect.
